---
title: "ManyBabies Simulated Analysis"
output:
html_document:
  toc: true
  theme: united
---

# Preliminaries

```{r Preliminaries}
options(dplyr.width = Inf)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache=TRUE)

library(ggplot2)
library(dplyr)
library(lme4)
library(tidyr)
library(magrittr)
library(lsmeans)
library(langcog)

theme_manylabs <- theme_bw() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, family="Arial"),
        text = element_text(family="Arial"),
        legend.key = element_rect(fill = "navy"),
        legend.background = element_rect(fill = "white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line.y = element_line(),
        axis.line.x = element_line(),
        strip.background = element_rect(fill = "white", colour = NA),
        strip.text.x = element_text(size = 14, face = "bold", colour = "black"))
```


# Simulate Data

Create basics of data.

```{r dataBasics}
data <- expand.grid(Lab = factor(paste0('Lab',1:20)),
                    Subject = 1:30,
                    Trial = 1:16)
```

Simulation sequence:

+ assign block
+ randomly assign conditions within blocks
+ randomly generate looking times within constraints
+ do it lab-by-lab and then add some subject attention diffs
+ for (1) realism and (2) convergence issues
+ max 20s looking per trial
+ randomly generate ages

*removed this from first batch and created a second batch varyng those two factors*
+ randomly generate method and participant vars (session, language, bilingual)

```{r mainData}
blocks <- factor(c('IDS','IDS','ADS','ADS'))
langs <- factor(c('American English','French','Spanish','German','Japanese', 'British English'))
methods <- factor(c('HPP','SingleScreen','AnotherMethod'))
sessions <- factor(c('First','Second'))
bilingual <- c(TRUE, FALSE)

lt_max <- 20        # max LT in a trial

data %<>%
  mutate(Subject = factor(toupper(paste0(Lab,'-',Subject))),
         Block = ((Trial-1) %/% 4) + 1) %>%
  group_by(Subject,Block) %>%
  mutate(Condition = sample(blocks)) %>%
  group_by(Lab) %>%
  mutate(.LabDiff = runif(1, 0, .5)) %>%
  ungroup() %>%
  mutate(LT = ifelse(Condition == 'IDS', 
                     rlnorm(n(), 1.5 + .LabDiff, .7), 
                     rlnorm(n(), 1.5, .7))) %>%
  group_by(Subject) %>%
  mutate(LT = LT + runif(1,0,.5)) %>%
  ungroup() %>%
  mutate(LT = ifelse(LT > lt_max, lt_max, LT)) %>%
  group_by(Lab) %>%
  mutate(.LabMeanAge = round(runif(1, 3, 12))) %>%
  ungroup() %>%
  group_by(Subject) %>%
  mutate(Age = round(runif(1, .LabMeanAge-.5, .LabMeanAge+.5),2)) %>%
  ungroup() %>%
  group_by(Lab) %>%
  mutate(Method = sample(methods,1)) %>%
  ungroup() %>%
  group_by(Subject) %>%
  mutate(Session = "First",
         Language = sample(langs, 1),
         Bilingual = FALSE) %>%
  arrange(Lab,Subject,Trial) %>%
  select(-starts_with("."))
```

Additional data labs contribute after testing the "promised" children is complete. 

+ Number of labs: 5 (instead of 20)
+ Lab-specific difference decreased (from 0-.5 to 0-.25)
+ Looking times slightly decreased

```{r additionalData}

additional_data <- expand.grid(Lab = factor(paste0('Lab',1:5)),
                    Subject = 1:30,
                    Trial = 1:16)

additional_data  %<>%
  mutate(Subject = factor(toupper(paste0(Lab,'-',Subject))),
         Block = ((Trial-1) %/% 4) + 1) %>%
  group_by(Subject,Block) %>%
  mutate(Condition = sample(blocks)) %>%
  group_by(Lab) %>%
  mutate(.LabDiff = runif(1, .0, .25)) %>%
  ungroup() %>%
  mutate(LT = ifelse(Condition == 'IDS', 
                     rlnorm(n(), 1.1 + .LabDiff, .8), 
                     rlnorm(n(), 1.1, .8))) %>%
  group_by(Subject) %>%
  mutate(LT = LT + runif(1,0,.5)) %>%
  ungroup() %>%
  mutate(LT = ifelse(LT > lt_max, lt_max, LT)) %>%
  group_by(Lab) %>%
  mutate(.LabMeanAge = round(runif(1, 3, 12))) %>%
  ungroup() %>%
  group_by(Subject) %>%
  mutate(Age = round(runif(1, .LabMeanAge-.5, .LabMeanAge+.5),2)) %>%
  ungroup() %>%
  group_by(Lab) %>%
  mutate(Method = sample(methods,1)) %>%
  ungroup() %>%
  group_by(Subject) %>%
  mutate(Session = sample(sessions,1),
         Language = sample(langs, 1),
         Bilingual = sample(bilingual, 1)) %>%
  arrange(Lab,Subject,Trial) %>%
  select(-starts_with("."))


```

# Data Cleaning

+ exclude outliers based on log-transformed LT
+ filter based on minimum trials of each type
+ remove hidden fields

```{r dataCleaning}
lt_min <- 2                # minumum LT for inclusion
z_threshold <- 3           # outlier threshold (sds)
min_trials_per_type <- 4   # min trials per type for inclusion

data_clean <- data %>%
  filter(LT >= lt_min) %>%
  group_by(Subject) %>%
  mutate(log_lt = log(LT), 
         .scaled_log_lt = as.numeric(langcog::scale(log_lt))) %>%
  filter(abs(.scaled_log_lt) < z_threshold) %>%
  group_by(Subject) %>%
  mutate(.N_IDS = sum(Condition == "IDS"),
         .N_ADS = sum(Condition == "ADS")) %>%
  filter(.N_IDS >= min_trials_per_type & 
           .N_ADS >= min_trials_per_type) %>%
  select(-starts_with("."))

additional_data_clean <- additional_data %>%
  filter(LT >= lt_min) %>%
  group_by(Subject) %>%
  mutate(log_lt = log(LT), 
         .scaled_log_lt = as.numeric(langcog::scale(log_lt))) %>%
  filter(abs(.scaled_log_lt) < z_threshold) %>%
  group_by(Subject) %>%
  mutate(.N_IDS = sum(Condition == "IDS"),
         .N_ADS = sum(Condition == "ADS")) %>%
  filter(.N_IDS >= min_trials_per_type & 
           .N_ADS >= min_trials_per_type) %>%
  select(-starts_with("."))
```

## Join two datasets for further analyses

```{r joinData}
all_data_clean = rbind(data_clean, additional_data_clean)
```

# Distributions 

"initial" data
```{r}
ggplot(data_clean, aes(x = LT)) + 
  geom_histogram() + 
  facet_grid(~Condition)
```

all data
```{r}
ggplot(all_data_clean, aes(x = LT)) + 
  geom_histogram() + 
  facet_grid(~Condition)
```

# Transformation

```{r}
ggplot(data_clean, aes(x=log_lt)) + 
  geom_histogram() + 
  facet_grid(~Condition)
```

# Create Aggregated Datasets

```{r}
agg_subjects <- data_clean %>%
  group_by(Lab, Method, Subject, Language, Condition, Age) %>%
  summarise(MeanLogLT = mean(log_lt)) %>%
  mutate(ConditionC = ifelse(Condition == "IDS", .5, -.5)) %>%
  mutate(Native = ifelse(Language == "American English", TRUE, FALSE))

agg_subjects_paired <- agg_subjects %>%
  select(-ConditionC) %>%
  spread(Condition, MeanLogLT) %>%
  mutate(Diff = IDS - ADS, 
         Prop = IDS / (IDS + ADS))

all_agg_subjects <- all_data_clean %>%
  group_by(Lab, Method, Session, Subject, Language, Bilingual, Condition, Age) %>%
  summarise(MeanLogLT = mean(log_lt)) %>%
  mutate(ConditionC = ifelse(Condition == "IDS", .5, -.5)) %>%
  mutate(Native = ifelse(Language == "American English", TRUE, FALSE))

all_agg_subjects_paired <- all_agg_subjects %>%
  select(-ConditionC) %>%
  spread(Condition, MeanLogLT) %>%
  mutate(Diff = IDS - ADS, 
         Prop = IDS / (IDS + ADS))

```

# Hypothesis Tests

## Overall Preference for IDS v ADS

```{r}
ggplot(agg_subjects, aes(x=Condition, y=MeanLogLT, group=Subject)) +
  geom_boxplot() +
  geom_jitter(width = .25, size =.5) +
  theme_manylabs

ggplot(agg_subjects_paired, aes(x='Overall', y=Diff)) +
  geom_boxplot() +
  geom_jitter(width = .25, size =.5) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

t.test(agg_subjects_paired$Diff)

mean(agg_subjects_paired$Diff) / sd(agg_subjects_paired$Diff)

model <- lmer(MeanLogLT ~ ConditionC + 
                (ConditionC | Lab) + 
                (1 | Subject), 
              data=agg_subjects, REML=FALSE)

summary(model)
drop1(model,~.,test="Chi")
```

### Lab Variability

```{r}
ggplot(agg_subjects_paired, aes(x=Lab, y=Diff)) +
  stat_summary(fun.y='mean', geom='point') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  coord_flip() +
  theme_manylabs

model <- lmer(MeanLogLT ~ ConditionC + (1 + ConditionC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
fixed_effect <- fixef(model)[['ConditionC']]

lab_ranefs <- data.frame(Lab = factor(rownames(ranef(model)$Lab), 
                                      levels=rownames(ranef(model)$Lab)),
                         ConditionRanef = ranef(model)$Lab$ConditionC + 
                           fixed_effect)

ggplot(lab_ranefs, aes(x=Lab, y=ConditionRanef, group=Lab)) +
  geom_point() +
  geom_errorbar(aes(ymin=fixed_effect, ymax=ConditionRanef), width=.1) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  geom_hline(yintercept=fixed_effect, linetype="solid", alpha=.5) +
  scale_x_discrete('') +
  scale_y_continuous('Random Effect') +
  coord_flip() +
  theme_manylabs
```

## Does IDS preference change by age?

```{r}
agg_subjects %<>%
  ungroup() %>%
  mutate(AgeC = Age - mean(Age))

ggplot(agg_subjects_paired, aes(x=Age, y=Diff)) +
  geom_point() +
  stat_smooth() +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  theme_manylabs

model <- lmer(MeanLogLT ~ ConditionC*AgeC + 
                (1 + ConditionC + AgeC | Lab) + 
                (1 | Subject), data=agg_subjects, 
              REML=FALSE)
summary(model)
drop1(model,~.,test="Chi")
```

### Does quadratic age term improve the fit?

```{r}
# model with linear+quadratic random effect of age, but only linear fixed effect
model <- lmer(MeanLogLT ~ ConditionC*poly(AgeC,1) + 
                (1 + ConditionC + poly(AgeC,2) | Lab) + 
                (1 | Subject), 
              data=agg_subjects, REML=FALSE)

# model with linear+quadratic random and fixed effects of age
model_2 <- lmer(MeanLogLT ~ ConditionC*poly(AgeC,2) + 
                  (1 + ConditionC + poly(AgeC,2) | Lab) + 
                  (1 | Subject), data=agg_subjects, 
                REML=FALSE)

anova(model,model_2)
```

## Are there trial order effects?

```{r}
data_clean %<>%
  ungroup() %>%
  mutate(TrialC = Trial - mean(Trial),
         ConditionC = ifelse(Condition == "IDS", .5, -.5),
         AgeC = Age - mean(Age))

ggplot(data_clean, aes(x=Trial, y=log_lt, color=Condition)) +
  stat_summary(fun.y='mean', geom='point') +
  stat_summary(fun.y='mean', geom='line') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  theme_manylabs

# model <- lmer(log_lt ~ ConditionC*AgeC*TrialC + 
#                 (1 + ConditionC + AgeC + TrialC | Lab) + 
#                 (1 + TrialC + ConditionC | Subject), 
#               data=data_clean, REML=FALSE)
# summary(model)
# drop1(model,~.,test="Chi")
```

## Moderator Analyses

### Method

Note `age | lab` doesn't converge.

```{r}
#NB: for some reason I cannot knit without adding this bit again. 
agg_subjects %<>%
  ungroup() %>%
  mutate(AgeC = Age - mean(Age))

ggplot(agg_subjects_paired, aes(x=Method, y=Diff)) +
  geom_boxplot() +
  geom_jitter(width = .25, size =.5) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Method) <- contr.sum(length(unique(agg_subjects$Method)))

model <- lmer(MeanLogLT ~ ConditionC * Method * AgeC + 
                (1 + ConditionC  | Lab) + 
                (1 | Subject), 
              data=agg_subjects, REML=FALSE)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + 
                     (1 + ConditionC + AgeC | Lab) + 
                     (1 | Subject), data=agg_subjects, 
                   REML=FALSE)

anova(model, model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Method, var="ConditionC", adjust="none")
```

### Session

Tested on all data, but very imbalanced. But: How to select subset?

```{r}
all_agg_subjects %<>%
  ungroup() %>%
  mutate(AgeC = Age - mean(Age))

ggplot(all_agg_subjects_paired, aes(x=Session, y=Diff)) +
  geom_boxplot()+
  geom_jitter(width = .25, size =.5)+
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

all_agg_subjects$Session = as.factor(all_agg_subjects$Session)
contrasts(all_agg_subjects$Session) <- contr.sum(length(unique(all_agg_subjects$Session)))

model <- lmer(MeanLogLT ~ ConditionC*Session*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=all_agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=all_agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Session, var="ConditionC", adjust="none")
```

### Nativeness of the test (American English versus everyone else)

Tested only on the "initial" dataset, analysis conditional on meeting a minimum N for each language to be included 



```{r Nativeness}
agg_subjects %<>%
  ungroup() %>%
  mutate(AgeC = Age - mean(Age))


ggplot(agg_subjects_paired, aes(x=Native, y=Diff)) +
  geom_boxplot()+
  geom_jitter(width = .25, size =.5)+
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Native) <- contr.sum(length(unique(agg_subjects$Native)))

model <- lmer(MeanLogLT ~ ConditionC*Native*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Native, var="ConditionC", adjust="none")
```

### Native Language
Tested only on the "initial" dataset, analysis conditional on meeting a minimum N for each language to be included 

```{r NativeLang}
ggplot(agg_subjects_paired, aes(x=Language, y=Diff)) +  
  geom_boxplot()+
  geom_jitter(width = .25, size =.5)+
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Language) <- contr.sum(length(unique(agg_subjects$Language)))

model <- lmer(MeanLogLT ~ ConditionC*Language*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Language, var="ConditionC", adjust="none")
```

### Biligualism
Tested on all data, but very imbalanced. But: How to select subset?

```{r}
ggplot(all_agg_subjects_paired, aes(x=Bilingual, y=Diff)) +
  geom_boxplot()+
  geom_jitter(width = .25, size =.5)+
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(all_agg_subjects$Bilingual) <- contr.sum(length(unique(all_agg_subjects$Bilingual)))

model <- lmer(MeanLogLT ~ ConditionC*Bilingual*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=all_agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=all_agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Bilingual, var="ConditionC", adjust="none")
```



