---
title: "ManyBabies Analysis"
output:
  html_document:
    toc: true
    theme: united
---

# Configuration

```{r}
z_threshold <- 2
min_trials_per_type <- 4
```

# Preliminaries

```{r}
options(dplyr.width = Inf)

library(ggplot2)
library(dplyr)
library(lme4)
library(tidyr)
library(lsmeans)
```

# Simulate Data

```{r}
data <- expand.grid(
              Lab = paste0('Lab',1:20),
              Subject = 1:30,
              Trial = 1:16
            )

data <- data %>%
          mutate(
            Lab = factor(Lab),
            Subject = toupper(paste0(Lab,'-',Subject)),
            Subject = factor(Subject)
          ) %>%
  
          # assign block
          mutate(
            Block = ((Trial-1) %/% 4) + 1
          ) %>%
  
          # randomly assign conditions within blocks
          group_by(Subject,Block) %>%
          mutate(
            Condition = sample(c('IDS','IDS','ADS','ADS'))
          ) %>%
          ungroup() %>%
          mutate(
            Condition = factor(Condition)
          ) %>%
  
          # randomly generate looking times within constraints
          # do it lab-by-lab and then add some subject attention diffs
          # for (1) realism and (2) convergence issues
          group_by(Lab) %>%
          mutate(
            .LabDiff = runif(1,0,.5)
          ) %>%
          ungroup() %>%
          mutate(
            LT = ifelse(Condition == 'IDS', rlnorm(n(),1.5+.LabDiff,.7), rlnorm(n(),1.5,.7)),
            LT = ifelse(LT > 20, 20, LT)
          ) %>%
          group_by(Subject) %>%
          mutate(
            LT = LT + runif(1,0,.5)
          ) %>%
  
          # randomly generate ages
          group_by(Lab) %>%
          mutate(
            .LabMeanAge = round(runif(1, 3,12))
          ) %>%
          ungroup() %>%
          group_by(Subject) %>%
          mutate(
            Age = round(runif(1, .LabMeanAge-.5, .LabMeanAge+.5),2)
          ) %>%
          ungroup() %>%
  
          # randomly generate method and participant vars (session, language, bilingual)
          group_by(Lab) %>%
          mutate(
            Method = sample(c('HPP','SingleScreen','AnotherMethod'),1)
          ) %>%
          ungroup() %>%
          group_by(Subject) %>%
          mutate(
            Session = sample(c('First','Second'),1),
            Language = sample(c('English','French','Spanish','German','Japanese'), 1),
            Bilingual = sample(c(1,0), 1)
          ) %>%
          ungroup() %>%
          mutate(
            Session = factor(Session),
            Method = factor(Method),
            Language = factor(Language),
            Bilingual = factor(Bilingual)
          ) %>%
    
          # arrange
          arrange(Lab,Subject,Trial) %>%
        
          # drop hidden fields  
          select(-starts_with("."))
```

# Data Cleaning

```{r}
data_clean <- data %>%
              filter(LT >= 2) %>%
            
              # exclude outliers based on log-transformed LT
              group_by(Subject) %>%
              mutate(
                .LogLT = log(LT),
                .SubjectMeanLT = mean(.LogLT)
              ) %>%
              ungroup() %>%
              mutate(
                .MeanLT = mean(.LogLT),
                .SDLT = sd(.LogLT)
              ) %>%
              mutate(
                .z = (.SubjectMeanLT - .MeanLT) / .SDLT
              ) %>%
              filter(
                abs(.z) < z_threshold
              ) %>%       
  
              # filter based on minimum trials of each type
              group_by(Subject) %>%
              mutate(
                .N_IDS = length(Condition[Condition == 'IDS']),
                .N_ADS = length(Condition[Condition == 'ADS'])
              ) %>%
              ungroup() %>%
              filter(
                .N_IDS >= min_trials_per_type & .N_ADS >= min_trials_per_type
              ) %>%

              # remove hidden fields
              select(-starts_with("."))
```

# Distributions

```{r}
ggplot(data_clean, aes(x=LT)) + geom_histogram() + facet_grid(~Condition)
```

# Transformation

```{r}
data_clean$LogLT <- log(data_clean$LT)

ggplot(data_clean, aes(x=LogLT)) + geom_histogram() + facet_grid(~Condition)
```

# Create Aggregated Datasets

```{r}
agg_subjects <- data_clean %>%
                group_by(Lab,Method,Session,Subject,Language,Bilingual,Condition,Age) %>%
                summarise(
                  MeanLogLT = mean(LogLT)
                ) %>%
                ungroup()

agg_subjects_paired <- agg_subjects %>%
                       spread(Condition,MeanLogLT) %>%
                       mutate(
                         Diff = IDS - ADS,
                         Prop = IDS / (IDS + ADS)
                       )
```

## ggplot Theme

```{r}
theme_manylabs <- theme_bw() +
                  theme(
                        axis.text = element_text(size = 14),
                        axis.title = element_text(size = 16, family="Arial"),
                        text = element_text(family="Arial"),
                        legend.key = element_rect(fill = "navy"),
                        legend.background = element_rect(fill = "white"),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        panel.border = element_blank(),
                        axis.line = element_line(),
                        strip.background = element_rect(fill = "white", colour = NA),
                        strip.text.x = element_text(size = 14, face = "bold", colour = "black")
                  )
```

# Hypothesis Tests

## Overall Preference for IDS v ADS

```{r}
ggplot(agg_subjects, aes(x=Condition, y=MeanLogLT)) +
  geom_boxplot() +
  theme_manylabs

ggplot(agg_subjects, aes(x=Condition, y=MeanLogLT)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  theme_manylabs
  
ggplot(agg_subjects_paired, aes(x='Overall', y=Diff)) +
  geom_boxplot() +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

ggplot(agg_subjects_paired, aes(x='Overall', y=Diff)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

t.test(agg_subjects_paired$Diff)

mean(agg_subjects_paired$Diff) / sd(agg_subjects_paired$Diff)

agg_subjects <- agg_subjects %>%
                mutate(
                  ConditionC = ifelse(Condition == 'IDS', .5, -.5),
                  ConditionC = ConditionC - mean(ConditionC)
                )

model <- lmer(MeanLogLT ~ ConditionC + (1 + ConditionC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)
drop1(model,~.,test="Chi")
```

### Lab Variability

```{r}
ggplot(agg_subjects_paired, aes(x=Lab, y=Diff)) +
              stat_summary(fun.y='mean', geom='point') +
              stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
              geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
              scale_x_discrete('') +
              coord_flip() +
              theme_manylabs

model <- lmer(MeanLogLT ~ ConditionC + (1 + ConditionC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
fixed_effect <- fixef(model)[['ConditionC']]

lab_ranefs <- data.frame(
                Lab = factor(rownames(ranef(model)$Lab), levels=rownames(ranef(model)$Lab)),
                ConditionRanef = ranef(model)$Lab$ConditionC + fixed_effect
            )

ggplot(lab_ranefs, aes(x=Lab, y=ConditionRanef, group=Lab)) +
        geom_point() +
        geom_errorbar(aes(ymin=fixed_effect, ymax=ConditionRanef), width=.1) +
        geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
        geom_hline(yintercept=fixed_effect, linetype="solid", alpha=.5) +
        scale_x_discrete('') +
        scale_y_continuous('Random Effect') +
        coord_flip() +
        theme_manylabs
```

## Does IDS preference change by age?

```{r}
ggplot(agg_subjects_paired, aes(x=Age, y=Diff)) +
  geom_point() +
  stat_smooth() +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  theme_manylabs

agg_subjects <- agg_subjects %>%
                mutate(
                  AgeC = Age - mean(Age)
                )

model <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)
drop1(model,~.,test="Chi")
```

### Does quadratic age term improve the fit?

```{r}
# model with linear+quadratic random effect of age, but only linear fixed effect
model <- lmer(MeanLogLT ~ ConditionC*poly(AgeC,1) + (1 + ConditionC + poly(AgeC,2) | Lab) + (1 | Subject), data=agg_subjects, REML=F)

# model with linear+quadratic random and fixed effects of age
model_2 <- lmer(MeanLogLT ~ ConditionC*poly(AgeC,2) + (1 + ConditionC + poly(AgeC,2) | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_2)
```

## Are there trial order effects?

```{r}
ggplot(data_clean, aes(x=Trial, y=LogLT, color=Condition)) +
  stat_summary(fun.y='mean', geom='point') +
  stat_summary(fun.y='mean', geom='line') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  theme_manylabs

data_clean <- data_clean %>%
                mutate(
                  TrialC = Trial - mean(Trial),
                  ConditionC = ifelse(Condition == 'IDS', .5, -.5),
                  ConditionC = ConditionC - mean(ConditionC),
                  AgeC = Age - mean(Age)
                )

model <- lmer(LogLT ~ ConditionC*AgeC*TrialC + (1 + ConditionC + AgeC + TrialC | Lab) + (1 + TrialC + ConditionC | Subject), data=data_clean, REML=F)
summary(model)
drop1(model,~.,test="Chi")
```

## Moderator Analyses

### Method

```{r}
ggplot(agg_subjects_paired, aes(x=Method, y=Diff)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Method) <- contr.sum(length(unique(agg_subjects$Method)))

model <- lmer(MeanLogLT ~ ConditionC*Method*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Method, var="ConditionC", adjust="none")
```

### Session

```{r}
ggplot(agg_subjects_paired, aes(x=Session, y=Diff)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Session) <- contr.sum(length(unique(agg_subjects$Session)))

model <- lmer(MeanLogLT ~ ConditionC*Session*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Session, var="ConditionC", adjust="none")
```

### Native Language

```{r}
ggplot(agg_subjects_paired, aes(x=Language, y=Diff)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Language) <- contr.sum(length(unique(agg_subjects$Language)))

model <- lmer(MeanLogLT ~ ConditionC*Language*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Language, var="ConditionC", adjust="none")
```

### Biligualism

```{r}
ggplot(agg_subjects_paired, aes(x=Bilingual, y=Diff)) +
  stat_summary(fun.y='mean', geom='bar') +
  stat_summary(fun.data='mean_cl_normal', geom='errorbar', width=.1, fun.args=list(mult=2)) +
  geom_hline(yintercept=0, linetype="dashed", alpha=.5) +
  scale_x_discrete('') +
  theme_manylabs

contrasts(agg_subjects$Bilingual) <- contr.sum(length(unique(agg_subjects$Bilingual)))

model <- lmer(MeanLogLT ~ ConditionC*Bilingual*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)
summary(model)

model_null <- lmer(MeanLogLT ~ ConditionC*AgeC + (1 + ConditionC + AgeC | Lab) + (1 | Subject), data=agg_subjects, REML=F)

anova(model,model_null)

# post-hoc least-squares contrasts
lstrends(model, ~ Bilingual, var="ConditionC", adjust="none")
```



