---
title: "ManyBabies 1: Pilot Data Analysis"
author: "Mike Frank, Hugh Rabagliati, and others"
output:
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
    highlight: tango
    theme: spacelab
---

This document shows an analysis of MB1 pilot data for purposes of procedural and analytic decision-making. It includes analysis of demographics and looking times, as well as condition differences. 

Please forgive errors, omissions, lack of clearly labeled axes, etc. This document was produced quickly to provide an initial guide for decision-making. 

There are many decision-points that need to be discussed throughout. Among them:

- What format to ask for data in? 
- What to do with minimum looking time trials (e.g., record as 2s, record as 0s, exclude)
- What trials to include in the analysis (e.g., should we include non-looking trials at the end of the experiemnt)?

```{r Preliminaries, results = "hide", message = FALSE}
options(dplyr.width = Inf)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache=FALSE)
library(lme4)
library(tidyverse)
library(stringr)
library(lubridate)
library(bit64) # necessary because of times from SMI > max integer size
library(langcog)
library(knitr)
library(forcats)
source("et_helper.R")

theme_set(theme_bw())
```

In general, based on the experiences below, we probably want to have three different spreadsheets:

- `labs` - this has lab and method info
- `subjects` - this has demographics for each subject
- `trials` - this has long-form trial data 

Each participating lab will contribute a `subjects` and a `trials` spreadsheet.

1. `trials` spreadsheet is a long-form data sheet. Each row is a trial. Columns are:

- `lab` [string] - unique identifier
- `subid` [string] - unique within-lab ID
- `age_days` [integer] - chronological age
- `trial_type` [string] - `IDS`, `ADS`, and `training`
- `stimulus` [string] - `training`, `IDS-X` or `ADS-x` - the actual sound file that was playing
- `trial_num` [integer] - trial number, from 1 -- 8 (with -2 and -1 denoting training trials)
- `looking_time` [double] - looking time in seconds
- `total_trial_time` [double] - total length of the trial

2. `subjects` spreadsheet is a long-form data sheet. Each row is a subject. Columns are:

- `lab` [string] - unique lab identifier
- `subid` [string] - unique ID 
- `method` [string] - `eye-tracking`, `HPP`, `single screen`


Then these can be merged appropriately. 


# Load data



## Frank data

Read in eye-tracker pilot data. 

```{r}
raw_data_path <- "pilot/frank/"
info_path <- "info/"
processed_data_path <- "processed_data/frank/"

all_data <- dir(raw_data_path, pattern="*.txt") %>%
  paste0(raw_data_path, .) %>%
  map_df(get_smi_header) %>% 
  split(.$file_name) %>%
  map_df(read_smi_idf) %>%
  split(.$file_name) %>%
  map_df(preprocess_data) 
```


Now extract trial numbers and match with stimuli.

Some items to be fixed:

* Currently we're not actually matching stimuli to trials (this will require parsing eye-tracker XML output)
* We're missing trial ADS-7; that's a (major) oversight from the experimental setup. 
* Right now I clip trials < 5, which trims out the two training trials. 

```{r}
frank_data <- all_data %>% 
  group_by(file_name, trial, stimulus) %>%
  summarise(looking_time = sum(dt[!is.na(x) & !is.na(y)]),
            looking_time = ifelse(is.na(looking_time), 0, looking_time)) %>%
  mutate(trial_cat = ifelse(str_detect(stimulus, ".jpg"), "speech","other")) %>%
  filter(trial_cat == "speech") %>%
  group_by(file_name) %>%
  filter(trial > 5) %>%
  mutate(trial_num = 1:n(), 
         subid = str_replace(str_replace(file_name,raw_data_path,""),
                           ".txt","")) 
```

Now merge in demographic information. 

```{r}
suppressWarnings(info <- read_csv("info/frank_demo.csv", col_types = cols()))

frank_data <- info %>% 
  select(subid, age, order) %>%
  left_join(frank_data)
```

Now merge in orders. 

```{r}
orders <- suppressWarnings(read_csv("info/orders.csv", col_types = cols())) %>%
  gather(marker, stimulus, 2:19) %>%
  rename(order = Order) %>%
  filter(!str_detect(stimulus, "Train")) %>% 
  group_by(order) %>%
  mutate(trial_num = 1:n()) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = -2) %>%
  select(-marker, -stim_num)

frank_data <- left_join(frank_data, orders) %>%
  mutate(trial_num = ceiling(trial_num  / 2)) %>%
  mutate(age_days = as.numeric(age), 
         lab = "stanford", 
         method = "eye-tracking") %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)

```

## Floccia data

```{r}
floccia_data <- suppressWarnings(read_csv("pilot/floccia/pilot data.csv", 
                                          col_types = cols())) %>%
  rename(age_days = age, 
         looking_time = LT) %>%
  mutate(subid = as.character(id), 
         method = "HPP",
         stimulus = str_replace(str_replace(stimulus, ".wav", ""), 
                                "Manybabies\\\\", "")) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = "-") %>%
  mutate(trial_num = ceiling(trial/2)) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
         
```

## Hamlin data

```{r}
hamlin_path <- "pilot/hamlin/"
hamlin_data <- dir(hamlin_path, pattern="*.csv") %>%
  paste0(hamlin_path, .) %>%
  map_df(function(x) {suppressWarnings(read_csv(x, col_types = cols())) %>% 
      mutate(order = x)}) %>%
  mutate(order = as.numeric(str_replace(str_replace(order, ".csv",""),
                                        "pilot/hamlin/order",""))) %>%
  gather(trial, looking_time, 
         starts_with("Train"), starts_with("IDS"), starts_with("ADS")) %>%
  separate(trial, into = c("trial_type","trial_num"), sep = -2) %>%
  mutate(lab = "ubc",
         method = "single-screen",
         trial_num = as.numeric(trial_num), 
         age_days = str_split(age, ";") %>% 
           map_dbl(function(x) as.numeric(x[1]) * 30.3 + as.numeric(x[2]))) %>%
  rename(subid = subnum) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
```

## Gonzalez-Gomez data

```{r}
gonzalez_data <- suppressWarnings(read_csv("pilot/gonzalez/long_data.csv", 
                                           col_types = cols())) %>%
  mutate(lab = "brookes", 
         method = "single-screen",
         subid = SubjectID, 
         age_days = as.numeric(str_replace(age_mo, "m", ""))*30.3,
         trial_type = ifelse(str_detect(StimName, "trial"), "Train",
                             ifelse(str_detect(StimName, "IDS"), "IDS", "ADS")), 
         trial_num = floor((Trial+1)/2), 
         looking_time = TotalLook/1000) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
```

## Merge all data 

This is what the eventual data frame looks like:

```{r}
d <- bind_rows(floccia_data, hamlin_data, frank_data, gonzalez_data)
kable(head(d))
```

# Descriptives and Exclusions

```{r}
d %>%
  group_by(lab, age_days, subid) %>%
  distinct %>%
  group_by(lab) %>%
  summarise(n = n(), 
            age_months = mean(age_days)/30.3) %>%
  kable(digits = 1)
```
## Exclusions

This is an important decision-point. At least we want to exclude:

- Children with no looking time
- Trials with no looking time

```{r}
lt_totals <- d %>%
  group_by(lab, subid) %>%
  summarise(total_looking_time = sum(looking_time, na.rm=TRUE))

d <- d %>%
  left_join(lt_totals) %>%
  filter(total_looking_time != 0, 
         !is.na(total_looking_time)) %>%
  mutate(looking_time = ifelse(looking_time == 0, NA, looking_time))
  

```

## Demographics

What's our participant distribution?

```{r}
subs <- d %>%
  group_by(lab, subid, age_days) %>%
  distinct

qplot(age_days, fill = lab, data=subs)
```

## Looking time dynamics

First, the overall distribution of looking times. 

```{r}
qplot(looking_time, fill = lab, facets = ~ lab, binwidth = 2, data = d)
```

Stanford has a large number of 2s looking times because that's the lookaway from the tracker. So when a child isn't looking at all, they get a 2s. *How should we deal with this?*

```{r}
qplot(age_days, looking_time, col = lab, data = d)
```

## Child outcomes

Next, are children making it through the experiment? Looks like essentially everyone does.

```{r}
d %>%
  group_by(lab, subid) %>%
  summarize(max_trial = max(trial_num[!is.na(looking_time)])) %>%
  summarise(prop_finishing = mean(max_trial == 8)) %>%
  kable(digits = 2)
```

Now, histogram of looking time by trial number. Looks like looking times are staying pretty long. 

```{r}
ggplot(d, aes(x = looking_time, fill = lab)) + 
  geom_histogram(binwidth = 2) + 
  facet_wrap(~trial_num)
```

We can look at this by age, too. 

```{r}
qplot(age_days, looking_time, col = lab, facets = ~ trial_num, data = d) + 
  geom_smooth(aes(group = 1), method = "lm", col = "black")
```

Plot means. Note that this graph has survivorship bias -- namely, those observations later in the graph represent kids that had more trials. 

```{r}
ms <- d %>%
  group_by(lab, trial_num) %>%
  multi_boot_standard(col = "looking_time", na.rm=TRUE)

ggplot(ms, aes(x = trial_num, y = mean, col = lab)) + 
  geom_line() + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width = .1)) 
```

We could plot this as a more classic habituation graph, from habituation.

```{r}
last_trial <- d %>%
  group_by(lab, subid) %>%
  summarise(max_trial = max(trial_num[!is.na(looking_time)])) 

ms <- d %>%
  left_join(last_trial) %>%
  mutate(trial_num_habit = trial_num - max_trial) %>%
  group_by(lab, trial_num_habit) %>%
  multi_boot_standard(col = "looking_time", na.rm=TRUE)

ggplot(ms, aes(x = trial_num_habit, y = mean, col = lab)) + 
  geom_smooth(se=FALSE, span = 1) +
  geom_point() + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width = .1)) + 
  xlab("Trials to habituation/experiment end") + 
  xlim(-7.1,0.1)
```

# Condition differences

For each pair of trials, subtract to get the difference score. Again following Csibra, we do a difference of logs. 

```{r}
diffs <- d %>%
  filter(trial_type != "Train", 
         looking_time != 0, 
         !is.na(looking_time)) %>%
  group_by(lab, subid, age_days, trial_num) %>%
  filter(sum(trial_type=="IDS") == 1 & 
           sum(trial_type=="ADS") == 1) %>% # take only pairs that are complete
  summarise(diff = looking_time[trial_type=="IDS"] - 
              looking_time[trial_type=="ADS"],
            diff_log = log10(looking_time[trial_type=="IDS"]) - 
              log10(looking_time[trial_type=="ADS"]))
```

What's the distributional form of these data? 

```{r}
qplot(diff, data = diffs) + 
  geom_vline(xintercept = mean(diffs$diff), col = "red", lty = 2) 
```

```{r}
qplot(diff_log, data = diffs) + 
  geom_vline(xintercept = mean(diffs$diff_log), col = "red", lty = 2) 
```

How do they change with trials?

```{r}
ms_diff <- diffs %>%
  group_by(lab, trial_num) %>%
  multi_boot_standard(col = "diff_log", na.rm=TRUE)

ggplot(ms_diff, aes(x = trial_num, y = mean)) +
         geom_smooth(se = FALSE, span = 2) + 
  facet_wrap(~lab) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width= .1)) +
  ylab("IDS preference (log10 s)") + 
  geom_hline(yintercept = 0, lty = 2)
```

Or with age? 

```{r}
qplot(age_days, diff_log, col = lab, group = 1, data = diffs) + 
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = 0, lty = 2) + 
  ylab("IDS preference (s)") 
```
         
By age and by trial. 

```{r}
qplot(age_days, diff_log, col = lab, group = 1, data = diffs) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~trial_num) + 
  geom_hline(yintercept = 0, lty = 2) + 
  ylab("IDS preference (s)") 
```


# Hypothesis tests

**I would very much recommend including random effects of item, but that info is not included with the data as is.**  

## Overall preference tests  
#### One Sample t-test against chance 
Using log transformed looking time.

```{r}
d_t_test1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time)) %>% 
  group_by(subid,trial_type) %>% 
  summarise(log_lt = mean(log_lt)) %>%
  group_by(subid) %>% 
  filter(n() == 2) %>%
  summarize(log_lt_diff = log_lt[trial_type == "IDS"] - log_lt[trial_type == "ADS"])
t.test(d_t_test1$log_lt_diff , mu = 0)
```

#### Paired t-test 

As above, but perhaps it is simpler to compare conditions, as our mixed effects models will use a condition predictor. Using log transformed looking time.
```{r}
d_t_test2 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time)) %>% 
  group_by(subid,trial_type) %>% 
  summarise(log_lt = mean(log_lt))
t.test(log_lt ~  trial_type, data = d_t_test2, paired = T)
```


## IDS Preference across age

#### Linear effect of age

Planned regression: 1 + CentredAge * Trial Type + (1 + Centred Age * Trial Type)  
I removed the by lab interaction to aid convergence.

```{r}
d_lmer1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days)) 
summary(lmer(log_lt ~ 1 + AgeC * trial_type + (1+ AgeC + trial_type|lab), data = d_lmer1))
```

#### Quadratic effect of age
Planned regression: 1 + (CentredAge + CentredAge^2) * Trial Type + (1 + (Centred Age + CentredAge ^2) * Trial Type)  
I removed the by lab interaction to aid convergence

```{r}
d_lmer2 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days)) 
summary(lmer(log_lt ~ 1 + poly(AgeC,2) * trial_type + (1+ poly(AgeC,2) + trial_type|lab), data = d_lmer2))
```

#### Secondary hypothesis tests, e.g trial order and age. 
We will fit a linear mixed effects model predicting all individual observations, with the structure:  
log(looking.time) ~ trial.num * stimulus * age + (trial.num * stimulus | subid) + (trial.num * stimulus * age | lab)  
NB. This is taken from the RRR. Does stimulus here refer to condition or item? I have taken it to refer to condition.  
Interactions removed to aid convergence.

```{r}
d_lmer3 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days),trial_numC = (trial_num - mean(trial_num))/sd(trial_num)) 
summary(lmer(log_lt ~ 1 + AgeC * trial_type * trial_numC + (1+ trial_type + trial_numC|subid)+ (1+ AgeC + trial_type + trial_numC|lab) , data = d_lmer3))
```



## Christina's Planned Moderator analyses
These analyses were planned to be by preference score, rather than by interactions. **Iff we are able to include item names in the data files, I would suggest going by interactions instead so that we can account for item effects with regard to, e.g., missing data.**    

#### Preference by method
Note that this analysis is perfectly confounded with lab at the moment, so it will not converge.   
ids.pref ~ method + (1 | lab)

```{r}
d_moder1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time),AgeC = (age_days - mean(age_days))/sd(age_days)) %>% 
  group_by(subid,trial_type,method,lab,AgeC) %>% 
  summarise(log_lt = mean(log_lt)) %>%
  group_by(subid,method,lab,AgeC) %>% 
  filter(n() == 2) %>%
  summarize(log_lt_diff = log_lt[trial_type == "IDS"] - log_lt[trial_type == "ADS"])

summary(lmer(log_lt_diff ~ method + (1|lab), data = d_moder1))
```

#### Preference by method interacting with age
Note that this analysis is perfectly confounded with lab at the moment, so it will not converge.   
ids.pref ~ method * age + (1 +age | lab)

```{r}

summary(lmer(log_lt_diff ~ method*AgeC + (1+AgeC|lab), data = d_moder1))
```

# Conclusions

Practical recommendations:

- Need to make standardized templates for `lab`, `subject`, and `trial` data. 
- Need to develop policies for data exclusion at the subject level (e.g., any child excluded)
- Need to walk through and select the planned analyses - this is going to be tricky!

Conclusions: It looks like we're seeing some IDS preference for each group, albeit at a different part of the experiment for each age/lab combo.
 