---
title: "ManyBabies 1: Pilot Data Analysis"
author: "Mike Frank (edits by Hugh Rabagliati)"
output:
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
    highlight: tango
    theme: spacelab
---

This document shows an analysis of MB1 pilot data for purposes of procedural and analytic decision-making. It includes analysis of demographics and looking times, as well as condition differences. *Condition differences (i.e., analyses of IDS preference) are included for purposes of analysis planning and should not be used for modification of procedure.*

Please forgive errors, omissions, lack of clearly labeled axes, etc. This document was produced quickly to provide an initial guide for decision-making. 

There are many decision-points that need to be discussed throughout. Among them:

- What format to ask for data in? 
- What to do with minimum looking time trials (e.g., record as 2s, record as 0s, exclude)
- What trials to include in the analysis (e.g., should we include non-looking trials at the end of the experiemnt)?

```{r Preliminaries, results = "hide", message = FALSE}
options(dplyr.width = Inf)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache=TRUE)
library(lme4)
library(tidyverse)
library(eyetrackingR)
library(stringr)
library(lubridate)
library(bit64) # necessary because of times from SMI > max integer size
library(langcog)
library(knitr)
library(forcats)
source("et_helper.R")

theme_set(theme_bw())
```

# Load data

Loading process to standardize pilot data from each lab. Current desideratum is a long-form data sheet with columns:

- `lab` [string] - unique identifier
- `method` [string] - `eye-tracking`, `HPP`, `single screen`
- `subid` [string] - unique within-lab ID
- `age_days` [integer] - chronological age
- `trial_type` [string] - `IDS`, `ADS`, and `training`
- `trial_num` [integer] - trial number, from 1 -- 8 (with -2 and -1 denoting training trials)
- `looking_time` [double] - looking time in seconds

In general, we probably want to have three different spreadsheets:

- `labs` - this has lab and method info
- `subjects` - this has demographics for each subject
- `trials` - this has long-form trial data 

Then these can be merged appropriately. 

_Questions_:

- Do we want order in here? Probably. 
- Do we want specific stimulus? Likely yes as well.


## Frank data

Read in eye-tracker pilot data. 

```{r}
raw_data_path <- "pilot/frank/"
info_path <- "info/"
processed_data_path <- "processed_data/frank/"

all_data <- dir(raw_data_path, pattern="*.txt") %>%
  paste0(raw_data_path, .) %>%
  map_df(get_smi_header) %>% 
  split(.$file_name) %>%
  map_df(read_smi_idf) %>%
  split(.$file_name) %>%
  map_df(preprocess_data) 
```

Now extract trial numbers and match with stimuli.

Some items to be fixed:

- Currently we're not actually matching stimuli to trials (this will require parsing eye-tracker XML output)
- We're missing trial ADS-7; that's a (major) oversight from the experimental setup. 
- Right now I clip trial < 5, which trims out the two training trials. 

```{r}
frank_data <- all_data %>% 
  group_by(file_name, trial, stimulus) %>%
  summarise(looking_time = max(t_stim)) %>%
  mutate(trial_cat = ifelse(str_detect(stimulus, ".jpg"), "speech","other")) %>%
  filter(trial_cat == "speech") %>%
  group_by(file_name) %>%
  filter(trial > 5) %>%
  mutate(trial_num = 1:n(), 
         subid = str_replace(str_replace(file_name,raw_data_path,""),
                           ".txt","")) 
```

Now merge in demographic information. 

```{r}
info <- read_csv("info/frank_demo.csv")

frank_data <- info %>% 
  select(subid, age, order) %>%
  left_join(frank_data)
```

Now merge in orders. 

```{r}
orders <- read_csv("info/orders.csv") %>%
  gather(marker, stimulus, 2:19) %>%
  rename(order = Order) %>%
  filter(!str_detect(stimulus, "Train")) %>% 
  group_by(order) %>%
  mutate(trial_num = 1:n()) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = -2) %>%
  select(-marker, -stim_num)

frank_data <- left_join(frank_data, orders) %>%
  mutate(trial_num = ceiling(trial_num  / 2)) %>%
  mutate(age_days = as.numeric(age), 
         lab = "stanford", 
         method = "eye-tracking") %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)

```

## Floccia data

```{r}
floccia_data <- read_csv("pilot/floccia/pilot data.csv") %>%
  rename(age_days = age, 
         looking_time = LT) %>%
  mutate(subid = as.character(id), 
         method = "HPP",
         stimulus = str_replace(str_replace(stimulus, ".wav", ""), 
                                "Manybabies\\\\", "")) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = "-") %>%
  mutate(trial_num = ceiling(trial/2)) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
         
```

## Hamlin data

```{r}
hamlin_path <- "pilot/hamlin/"
hamlin_data <- dir(hamlin_path, pattern="*.csv") %>%
  paste0(hamlin_path, .) %>%
  map_df(function(x) {read_csv(x) %>% mutate(order = x)}) %>%
  mutate(order = as.numeric(str_replace(str_replace(order, ".csv",""),
                                        "pilot/hamlin/order",""))) %>%
  gather(trial, looking_time, 
         starts_with("Train"), starts_with("IDS"), starts_with("ADS")) %>%
  separate(trial, into = c("trial_type","trial_num"), sep = -2) %>%
  mutate(lab = "ubc",
         method = "single-screen",
         trial_num = as.numeric(trial_num), 
         age_days = str_split(age, ";") %>% 
           map_dbl(function(x) as.numeric(x[1]) * 30.3 + as.numeric(x[2]))) %>%
  rename(subid = subnum) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
```

## Merge all data 

This is what the eventual data frame looks like:

```{r}
d <- bind_rows(floccia_data, hamlin_data, frank_data)
kable(head(d))
```

## Descriptives

```{r}
d %>%
  group_by(lab, age_days, subid) %>%
  distinct %>%
  group_by(lab) %>%
  summarise(n = n(), 
            age_months = mean(age_days)/30.3) %>%
  kable(digits = 1)
```

# Plots

## Demographics

What's our participant distribution?

```{r}
subs <- d %>%
  group_by(lab, subid, age_days) %>%
  distinct

qplot(age_days, fill = lab, data=subs)
```

## Looking time dynamics

First, the overall distribution of looking times. 

```{r}
qplot(looking_time, fill = lab, facets = ~ lab, binwidth = 2, data = d)
```

Stanford has a large number of 2s looking times because that's the lookaway from the tracker. So when a child isn't looking at all, they get a 2s. *How should we deal with this?*

Next, are children making it through the experiment? Looks like essentially everyone does.

```{r}
final_trial <- d %>%
  group_by(lab, subid) %>%
  summarize(max_trial = max(trial_num[looking_time > 2]))
                            
qplot(max_trial, fill = lab, data = final_trial)
```

Now, histogram of looking time by trial number. Looks like looking times are staying pretty long. 

```{r}
ggplot(d, aes(x = looking_time, fill = lab)) + 
  geom_histogram(binwidth = 2) + 
  facet_wrap(~trial_num)
```

We can look at this by age, too. 

```{r}
qplot(age_days, looking_time, col = lab, facets = ~ trial_num, data = d) + 
  geom_smooth(aes(group = 1), method = "lm", col = "black")
```

Plot means. 

```{r}
ms <- d %>%
  group_by(lab, trial_num) %>%
  multi_boot_standard(col = "looking_time", na.rm=TRUE)

ggplot(ms, aes(x = trial_num, y = mean, col = lab)) + 
  geom_line() + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width = .1)) 
```

## Condition differences

### Between-subjects analysis

Doesn't compute difference scores - treats IDS and ADS observations as independent. These analyses should likely be lower-powered. (Interested to hear people's thoughts as to whether they should be included). 

```{r}
ms <- d %>%
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%  
  group_by(trial_num, trial_type) %>%
  multi_boot_standard(col = "looking_time", na.rm=TRUE)

ggplot(ms, aes(x = trial_num, y = mean, col = trial_type)) + 
  geom_line() + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width= .1)) +
  ylab("Looking time (s)")
```

Split by lab (which is really age now).

```{r}
ms <- d %>%
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%  
  group_by(lab, trial_num, trial_type) %>%
  multi_boot_standard(col = "looking_time", na.rm=TRUE) %>%
  ungroup %>%
  mutate(lab = fct_relevel(lab, "ubc"))

ggplot(ms, aes(x = trial_num, y = mean, col = trial_type)) + 
  geom_smooth(se = FALSE, span = 2) + 
  facet_wrap(~lab) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width= .1))+ 
  ylab("Looking time (s)") + 
  scale_color_solarized(name = "Trial Type") + 
  theme(legend.position = "bottom")
```

Take a look at this in log space as well, following Csibra et al. (2015), "Statistical Treatment of Looking-Time Data." Doesn't change much, but is likely better. 

```{r}
ms <- d %>%
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  group_by(lab, trial_num, trial_type) %>%
  mutate(log10_looking_time = log10(looking_time)) %>%
  multi_boot_standard(col = "log10_looking_time", na.rm=TRUE)

ggplot(ms, aes(x = trial_num, y = mean, col = trial_type)) + 
  geom_smooth(se = FALSE, span = 2) + 
  facet_wrap(~lab) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width= .1)) +
  ylab("Log looking time (s)") + 
  scale_color_solarized(name = "Trial Type") + 
  theme(legend.position = "bottom")
  
```

### Within-subjects

Now do (perhaps) the more appropriate analysis: For each pair of trials, subtract to get the difference score. Again following Csibra, we do a difference of logs. 

```{r}
diffs <- d %>%
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  group_by(lab, subid, age_days, trial_num) %>%
  filter(n() == 2) %>% # take only pairs that are complete
  summarise(idspref = log10(looking_time[trial_type=="IDS"]) - 
              log10(looking_time[trial_type=="ADS"]))
```

What's the distributional form of these data? 

```{r}
qplot(idspref, data = diffs)
```

How do they change with trials?

```{r}
ms_diff <- diffs %>%
  group_by(lab, trial_num) %>%
  multi_boot_standard(col = "idspref", na.rm=TRUE)

ggplot(ms_diff, aes(x = trial_num, y = mean)) +
         geom_smooth(se = FALSE, span = 2) + 
  facet_wrap(~lab) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position = position_dodge(width= .1)) +
  ylab("IDS preference (log10 s)") + 
  geom_hline(yintercept = 0, lty = 2)
```

Or with age? 

```{r}
qplot(age_days, idspref, col = lab, group = 1, data = diffs) + 
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = 0, lty = 2) + 
  ylab("IDS preference (s)") 
```
         
By age and by trial. 

```{r}
qplot(age_days, idspref, col = lab, group = 1, data = diffs) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~trial_num) + 
  geom_hline(yintercept = 0, lty = 2) + 
  ylab("IDS preference (s)") 
```


# Hypothesis tests
**I would very much recommend including random effects of item, but that info is not included with the data as is.**  

## Overall preference tests  
#### One Sample t-test against chance 
Using log transformed looking time.
```{r}
d_t_test1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time)) %>% 
  group_by(subid,trial_type) %>% 
  summarise(log_lt = mean(log_lt)) %>%
  group_by(subid) %>% 
  filter(n() == 2) %>%
  summarize(log_lt_diff = log_lt[trial_type == "IDS"] - log_lt[trial_type == "ADS"])
t.test(d_t_test1$log_lt_diff , mu = 0)
```

#### Paired t-test 
As above, but perhaps it is simpler to compare conditions, as our mixed effects models will use a condition predictor. Using log transformed looking time.
```{r}
d_t_test2 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time)) %>% 
  group_by(subid,trial_type) %>% 
  summarise(log_lt = mean(log_lt))
t.test(log_lt ~  trial_type, data = d_t_test2, paired = T)
```


## IDS Preference across age
#### Linear effect of age
Planned regression: 1 + CentredAge * Trial Type + (1 + Centred Age * Trial Type)  
I removed the by lab interaction to aid convergence.
```{r}
d_lmer1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days)) 
summary(lmer(log_lt ~ 1 + AgeC * trial_type + (1+ AgeC + trial_type|lab), data = d_lmer1))
```

#### Quadratic effect of age
Planned regression: 1 + (CentredAge + CentredAge^2) * Trial Type + (1 + (Centred Age + CentredAge ^2) * Trial Type)  
I removed the by lab interaction to aid convergence

```{r}
d_lmer2 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days)) 
summary(lmer(log_lt ~ 1 + poly(AgeC,2) * trial_type + (1+ poly(AgeC,2) + trial_type|lab), data = d_lmer2))
```

#### Secondary hypothesis tests, e.g trial order and age. 
We will fit a linear mixed effects model predicting all individual observations, with the structure:  
log(looking.time) ~ trial.num * stimulus * age + (trial.num * stimulus | subid) + (trial.num * stimulus * age | lab)  
NB. This is taken from the RRR. Does stimulus here refer to condition or item? I have taken it to refer to condition.  
Interactions removed to aid convergence.

```{r}
d_lmer3 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time), AgeC = (age_days - mean(age_days))/sd(age_days),trial_numC = (trial_num - mean(trial_num))/sd(trial_num)) 
summary(lmer(log_lt ~ 1 + AgeC * trial_type * trial_numC + (1+ trial_type + trial_numC|subid)+ (1+ AgeC + trial_type + trial_numC|lab) , data = d_lmer3))
```



## Christina's Planned Moderator analyses
These analyses were planned to be by preference score, rather than by interactions. **Iff we are able to include item names in the data files, I would suggest going by interactions instead so that we can account for item effects with regard to, e.g., missing data.**    

#### Preference by method
Note that this analysis is perfectly confounded with lab at the moment, so it will not converge.   
ids.pref ~ method + (1 | lab)

```{r}
d_moder1 <- d %>% 
  filter(trial_type != "Train", 
         looking_time != 0, !is.na(looking_time)) %>%
  mutate(log_lt = log(looking_time),AgeC = (age_days - mean(age_days))/sd(age_days)) %>% 
  group_by(subid,trial_type,method,lab,AgeC) %>% 
  summarise(log_lt = mean(log_lt)) %>%
  group_by(subid,method,lab,AgeC) %>% 
  filter(n() == 2) %>%
  summarize(log_lt_diff = log_lt[trial_type == "IDS"] - log_lt[trial_type == "ADS"])

summary(lmer(log_lt_diff ~ method + (1|lab), data = d_moder1))
```

#### Preference by method interacting with age
Note that this analysis is perfectly confounded with lab at the moment, so it will not converge.   
ids.pref ~ method * age + (1 +age | lab)

```{r}

summary(lmer(log_lt_diff ~ method*AgeC + (1+AgeC|lab), data = d_moder1))
```

# Conclusions

Practical recommendations:

- Need to make standardized templates for `lab`, `subject`, and `trial` data. 
- Need to develop policies for data exclusion at the subject level (e.g., any child excluded)
- Need to walk through and select the planned analyses - this is going to be tricky!

Conclusions: It looks like we're seeing some IDS preference for each group, albeit at a different part of the experiment for each age/lab combo.
 