---
title: "ManyBabies 1: Data Munging"
author: "ManyBabies Team"
output:
html_document:
toc: true
number_sections: true
code_folding: hide
highlight: tango
theme: spacelab
---

This script will munge all data and produce an output file for processing.

```{r Preliminaries, results = "hide", message = FALSE}
options(dplyr.width = Inf)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache=FALSE)
library(lme4)
library(tidyverse)
library(stringr)
library(lubridate)
library(bit64) # necessary because of times from SMI > max integer size
library(langcog)
library(knitr)
library(forcats)
source("et_helper.R")

theme_set(theme_bw())
```

## Frank data

Read in eye-tracker pilot data. 

```{r}
raw_data_path <- "pilot/frank/"
info_path <- "info/"
processed_data_path <- "processed_data/frank/"

all_data <- dir(raw_data_path, pattern="*.txt") %>%
  paste0(raw_data_path, .) %>%
  map_df(get_smi_header) %>% 
  split(.$file_name) %>%
  map_df(read_smi_idf) %>%
  split(.$file_name) %>%
  map_df(preprocess_data) 
```


Now extract trial numbers and match with stimuli.

Some items to be fixed:

* Currently we're not actually matching stimuli to trials (this will require parsing eye-tracker XML output)
* We're missing trial ADS-7; that's a (major) oversight from the experimental setup. 
* Right now I clip trials < 5, which trims out the two training trials. 

```{r}
frank_data <- all_data %>% 
  group_by(file_name, trial, stimulus) %>%
  summarise(looking_time = sum(dt[!is.na(x) & !is.na(y)]),
            looking_time = ifelse(is.na(looking_time), 0, looking_time)) %>%
  mutate(trial_cat = ifelse(str_detect(stimulus, ".jpg"), "speech","other")) %>%
  filter(trial_cat == "speech") %>%
  group_by(file_name) %>%
  filter(trial > 5) %>%
  mutate(trial_num = 1:n(), 
         subid = str_replace(str_replace(file_name,raw_data_path,""),
                             ".txt","")) 
```

Now merge in demographic information. 

```{r}
suppressWarnings(info <- read_csv("info/frank_demo.csv", col_types = cols()))

frank_data <- info %>% 
  select(subid, age, order) %>%
  left_join(frank_data)
```

Now merge in orders. 

```{r}
orders <- suppressWarnings(read_csv("info/orders.csv", col_types = cols())) %>%
  gather(marker, stimulus, 2:19) %>%
  rename(order = Order) %>%
  filter(!str_detect(stimulus, "Train")) %>% 
  group_by(order) %>%
  mutate(trial_num = 1:n()) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = -2) %>%
  select(-marker, -stim_num)

frank_data <- left_join(frank_data, orders) %>%
  mutate(trial_num = ceiling(trial_num  / 2)) %>%
  mutate(age_days = as.numeric(age), 
         lab = "stanford", 
         method = "eye-tracking") %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)

```

## Floccia data

```{r}
floccia_data <- suppressWarnings(read_csv("pilot/floccia/pilot data.csv", 
                                          col_types = cols())) %>%
  rename(age_days = age, 
         looking_time = LT) %>%
  mutate(subid = as.character(id), 
         method = "HPP",
         stimulus = str_replace(str_replace(stimulus, ".wav", ""), 
                                "Manybabies\\\\", "")) %>%
  separate(stimulus, into = c("trial_type", "stim_num"), sep = "-") %>%
  mutate(trial_num = ceiling(trial/2)) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)

```

## Hamlin data

```{r}
hamlin_path <- "pilot/hamlin/"
hamlin_data <- dir(hamlin_path, pattern="*.csv") %>%
  paste0(hamlin_path, .) %>%
  map_df(function(x) {suppressWarnings(read_csv(x, col_types = cols())) %>% 
      mutate(order = x)}) %>%
  mutate(order = as.numeric(str_replace(str_replace(order, ".csv",""),
                                        "pilot/hamlin/order",""))) %>%
  gather(trial, looking_time, 
         starts_with("Train"), starts_with("IDS"), starts_with("ADS")) %>%
  separate(trial, into = c("trial_type","trial_num"), sep = -2) %>%
  mutate(lab = "ubc",
         method = "single-screen",
         trial_num = as.numeric(trial_num), 
         age_days = str_split(age, ";") %>% 
           map_dbl(function(x) as.numeric(x[1]) * 30.3 + as.numeric(x[2]))) %>%
  rename(subid = subnum) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
```

## Gonzalez-Gomez data

Looks - LooksAway = LookingTime, per email from NGG. 

```{r}
gonzalez_data <- suppressWarnings(read_csv("pilot/gonzalez/long_data.csv", 
                                           col_types = cols())) %>%
  mutate(lab = "brookes", 
         method = "single-screen",
         subid = SubjectID, 
         age_days = as.numeric(str_replace(age_mo, "m", ""))*30.3,
         trial_type = ifelse(str_detect(StimName, "Trial"), "Train",
                             ifelse(str_detect(StimName, "IDS"), "IDS", "ADS")), 
         trial_num = floor((Trial+1)/2), 
         looking_time = (TotalLook - TotalLookAway)/1000) %>%
  select(lab, method, subid, age_days, trial_type, trial_num, looking_time)
```

## Merge all data 

This is what the eventual data frame looks like:

```{r}
d <- bind_rows(floccia_data, hamlin_data, frank_data, gonzalez_data)
kable(head(d))
```

```{r}
write_csv(d, "processed_data/pilot_data.csv")
```