---
title: "new ES calculations"
author: "Alejandrina Cristia"
date: "6/9/2017"
output:
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(metafor)
library(pwr)

```

## Read in data & calc ES



```{r data}
read.csv("ids_ma.csv")->x
table(x$coder,x$participant_design)
#new entries don't have ES, they are all within_two
summary(x$corr) #all NA

x$pooled_SD=x$corr_imp=x$d_calc=x$d_calc_var=x$es_method=NA
x$corr_imp=0

#between calc
x$pooled_SD[x$participant_design=="between"]<- sqrt(((x$n_1[x$participant_design=="between"] - 1) * x$SD_1[x$participant_design=="between"] ^ 2 + (x$n_2[x$participant_design=="between"] - 1) * x$SD_2[x$participant_design=="between"] ^ 2) / (x$n_1[x$participant_design=="between"] + x$n_2[x$participant_design=="between"] - 2)) # Lipsey & Wilson, 3.14
x$es_method[x$participant_design=="between"]  <- "between"

#within_two calc
#summary(x[x$participant_design=="within_two",c("x_1","x_2","SD_1","SD_2")]) #checking - 41 NAs corresponds to the original studies
x$pooled_SD[x$participant_design=="within_two"]<- sqrt((x$SD_1[x$participant_design=="within_two"] ^ 2 + x$SD_2[x$participant_design=="within_two"] ^ 2) / 2) # Lipsey & Wilson, 3.14
x$es_method[x$participant_design=="within_two"]  <- "group_means_two"

x$d_calc <- (x$x_1 - x$x_2) / x$pooled_SD # Lipsey & Wilson (2001)
x$d_var_calc <- (2 * (1 - x$corr_imp)/ x$n_1) + (x$d_calc ^ 2 / (2 *x$n_1)) # Lipsey & Wilson (2001)

summary(x$d_var_calc) #50 NAs
x$d_calc[x$coder=="Alex Cristia"]<-x$d[x$coder=="Alex Cristia"]
x$d_var_calc[x$coder=="Alex Cristia"]<-x$d_var[x$coder=="Alex Cristia"]
summary(x$d_var_calc) #no NAs

x$se=sqrt(x$d_var_calc)
x$w=1/sqrt(x$d_var_calc)
```

## Fit MA: All data

This is the MA on the Dunst data

```{r original}
original=rma.uni(yi=d_calc, sei=se, weights=w,data=x,subset=c(coder=="Alex Cristia"))
summary(original)
```

Next, we consider in addition papers entered by the MetaLab team

```{r all}
withNew=rma.uni(yi=d_calc, sei=se, weights=w,data=x)
summary(withNew)
```

## Contents of MA table

This is a rather diverse dataset, as the following levels portray.

```{r describe}
table(x$infant_type)
table(x$stim_language)
table(x$setting)
table(x$speaker)
table(x$speech_type)
table(x$dependent_measure)
```

## Subset analyses: Most relevant

Focus on:

- looking times collected in a laboratory 
- typically-developing infants 
- aged 3???15 months 
- natural IDS in English 
- from an unfamiliar female speaker

```{r mostrelevant}
subset(x, speech_type != "Filtered" & speech_type != "Synthesized" &
         speaker!="Child\342\200\231s mother" & speaker != "Unfamiliar male" & 
         setting=="Laboratory" & 
         stim_language=="English" & 
         infant_type=="typical" & 
         dependent_measure=="looking_time" &
         mean_age_1>=3*30.25 & mean_age_1<=15*30.25)->relevant
selected=rma.uni(yi=d_calc, sei=se, weights=w,data=relevant)
summary(selected)
```

And with this estimate we can do a power ana for single studies.

```{r pwr-calc}
pwr.t.test(d=selected$b[1],n=16,sig.level=.05,type="paired",alternative="greater")

```

## Subset analyses: Split by age

Important: in our relevant subset, age is not a significant moderator.

```{r withage}
relevant$age.c=relevant$mean_age_1-mean(relevant$mean_age_1,na.rm=T)
with_age=rma.uni(yi=d_calc, sei=se, weights=w,mods=age.c,data=relevant)
summary(with_age)
```

However, to be informative, next I show results for age groups relevant to ManyBabies:

```{r splitage}
for(thisage in c(3,6,9,12)){
  subset(relevant,  mean_age_1>=thisage*30.25 & mean_age_1< (thisage+3) *30.25)->agesub
  print(paste("from", thisage, "to", thisage+3))
  if(dim(agesub)[1]>2) print(summary(rma.uni(yi=d_calc, sei=se, weights=w,data=agesub))) else print("not enough studies")
}


```